Parsing Postfix log files

### 2007-05-29 doc/logparser.tex
Added cached vs discarded graphs for regex run time. Added Makefile rules to create the graphs. Improved related paper section.

    \item The first matching rule wins: no further rules are tried against
        that line,\footnote{With one exception, details are provided later
        in Section~\ref{actions-in-detail}.} but there is a facility for
        the specifying the order of the rules so that more specific rules
        can be tried first.

@@

    \item [regex] The regular expression\footnote{See
        http://en.wikipedia.org/wiki/Regular\_expression for more
        information about regular expressions in general.} to match the log
        line against.  The regex will first have several keywords expanded:
        this simplifies reading and writing rules, avoids needless
        repetition of complex regex components, and allows the components
        to be corrected and/or improved in one location.

        The following keywords are expanded:

        \texttt{\_\_SENDER\_\_, \_\_RECIPIENT\_\_, \_\_HELO\_\_,
        \_\_HOSTNAME\_\_, \_\_IP\_\_, \newline\_\_SMTP\_CODE\_\_,
        \_\_QUEUEID\_\_ and \_\_COMMAND\_\_.}

        \_\_COMMAND\_\_ is an SMTP command, e.g. MAIL FROM\@; hopefully the
        others are reasonably self explanatory.

        For efficiency the keywords are expanded and every rule's regex is
        compiled before attempting to parse the log file --- otherwise each
        regex would be recompiled each time it was used, resulting in an
        468.95\% slowdown.

            \item [Compiled once and saved:] run time was 55.47 seconds,
                standard deviation was 0.247 seconds.

            \item [Compiled every time and discarded:] run time was 260.14
                seconds (4 minutes, 20.14 seconds) (468.95\% slower),
                standard deviation was 0.96 seconds.

        XXX DO I NEED TO SAY ANYTHING MORE ABOUT THIS??

@@

>>>>

    \item The first matching rule wins: no further rules are tried against
        that line, but there is a facility for specifying the order of
        rules so that more specific rules can be tried first.

@@

    \item [regex] The regular expression\footnote{See
        http://en.wikipedia.org/wiki/Regular\_expression for more
        information about regular expressions in general.} to match the log
        line against.  The regex will first have several keywords expanded:
        this simplifies reading and writing rules, avoids needless
        repetition of complex regex components, allows the components
        to be corrected and/or improved in one location, and makes each
        regex largely self-documenting.

        The following keywords are expanded (full explanations can be found
        in the source code):

        \_\_SENDER\_\_, \_\_RECIPIENT\_\_, \_\_MESSAGE\_ID\_\_,
        \_\_HELO\_\_, \newline \_\_EMAIL\_\_, \_\_HOSTNAME\_\_, \_\_IP\_\_,
        \_\_IPv4\_\_, \_\_IPv6\_\_, \newline \_\_SMTP\_CODE\_\_,
        \_\_RESTRICTION\_START\_\_, \_\_QUEUEID\_\_, \newline
        \_\_COMMAND\_\_, \_\_SHORT\_CMD\_\_, \_\_DELAYS\_\_, \_\_DELAY\_\_,
        \_\_DSN\_\_ and \_\_CONN\_USE\_\_.

        For efficiency the keywords are expanded and every rule's regex is
        compiled before attempting to parse the log file --- otherwise each
        regex would be recompiled each time it was used, resulting in a
        large, data dependent slowdown.  Graph \ref{normal regex vs discard
        regex} shows execution times gathered over 93 log files, with and
        without caching the regex; also shown on the graph are the file
        sizes and number of lines in each file.  Graph \ref{normal regex vs
        disarded regex factor} shows the percentage execution time increase
        when not caching each regex.  Logging to the database was disabled
        for the test runs.

@@

\section{Graphs}

\subsection{Regex: cached vs compiled every time}
\label{normal regex vs discard regex}
\includegraphics{plot-time.ps}

\subsection{Regex: percentage execution time increase}
\label{normal regex vs disarded regex factor}
\includegraphics{plot-factor.ps}
